---
title: "Analysis Script"
author: "Sophia Drewry"
date: "10/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script loads the cleaned and processed data to perform some formal statistical fitting
```{r}
# load needed packages. make sure they are installed.
library(dplyr) #for data processing
library(here) #to set paths
library(tidymodels) #to fit models
library(dotwhisker) #for box and ehiskers plots
library(ggplot2)
```
#Load data
```{r}
# note the use of the here() package and not absolute paths
dataSPOT <- here::here("data","processed_data","processeddta.rds")
dataSPOT2 <- here::here("data","processed_data","feverdta.rds")
# load data. 
# note. processeddta is our main cleaned fata set. Fever data is just those who have fever, 
processeddta <-readRDS(dataSPOT)
feverdta <-readRDS(dataSPOT2)
# take a look at the main data
dplyr::glimpse(processeddta)
```
### End goal: 
- Main continuous outcome of interest = Body temperature
- Main categorical outcome = Nausea
- Main predictor of interest = Runny Nose
- We want to see if the other symptoms are correlated with (predict) those outcomes we looked at in the EDA.
The other predictors of interest are as follows:
- Fatigue (y/n)
- CoughIntensity (none/mild/moderate/severe)
- ChillsSweats (y/n)
- SwollenLymphNodes (y/n)
- Pharyngitis (y/n)
- Insomnia (y/n)
- Headache (y/n)
- SubjectiveFever (y/n)

# Model 1. Fitting a linear model to BodyTemp using RunnyNose the main predictor of interest
```{r}
# Using the tidymodels package, I am going to set up the linear model
# Set the engine
lm_mod <-  linear_reg() %>% 
  set_engine("lm") 
# Now I am fitting the model
lm_fit <- 
  lm_mod %>% 
  fit(BodyTemp ~ RunnyNose, data = processeddta)
tidy(lm_fit)
```
### Model as box and whiskers plot
```{r}
tidy(lm_fit) %>% 
dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))

```
# Model 2. Fitting a linear model to BodyTemp using all relevant predictors of interest
If we remember from our EDA, the other interesting predictors were

- CoughIntensity (none/mild/moderate/severe)
- ChillsSweats (y/n)
- SubjectiveFever (y/n)

*Note, I am not using Pharyngitis because if did not offer an interesting relationship in both body temperature variable alone as well as fever variable.
```{r}
# Using the tidymodels package, I am going to set up the linear model
# Set the engine
lm_mod <-  linear_reg() %>% 
  set_engine("lm") 
# Now I am fitting the model
lm_fit2 <- 
  lm_mod %>% 
  fit(BodyTemp ~ RunnyNose * ChillsSweats * CoughIntensity * SubjectiveFever, data = processeddta)
tidy(lm_fit2)

```
### Model as box and whiskers plot
```{r}
tidy(lm_fit2) %>% 
dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))

```

# Comparing Model 1 and Model 2
To best compare these models and see if they are significant, lets use an ANOVA table
```{r}
tidy(lm_fit)
tidy(lm_fit2)

ANOVA1<- anova(lm_fit$fit, lm_fit2$fit, test = "Chisq")
ANOVA1

```
P value is significant, on the more complex model. p<0.001

# Predicting fever temps using two different datasets
### Create model 2.2 with fever data set
```{r}
# First lets create a model with feverdta
lm_mod <-  linear_reg() %>% 
  set_engine("lm") 
# Now I am fitting the model
lm_fit2.2 <- 
  lm_mod %>% 
  fit(BodyTemp ~ RunnyNose * ChillsSweats * CoughIntensity * SubjectiveFever, data = feverdta)
tidy(lm_fit2.2)
```
### Setting Predictions 
```{r}
# Setting what we want to predict, a fever
new_points <- expand.grid(BodyTemp = 100.4,
                          CoughIntensity = c("None","Mild", "Moderate", "Severe"))
new_points
```
### Set prediction for model 2
```{r}

mean_pred <- predict(lm_fit2$fit, new_data = new_points)
#Set Confidence Interval
confint <- predict(lm_fit2$fit, 
                         new_data = new_points,
                         interval = 'confidence')
head(confint)
tail(confint)

#For some reason I am unable to bind mean_pred to confint for plotting and futher comparison
# So I am trowing in the towel on this one, but comparing outputs are interesting

```
### Set prediction for model 2.2
```{r}
mean_pred2 <- predict(lm_fit2.2$fit, new_data = new_points)
#Set Confidence Interval
confint2 <- predict(lm_fit2.2$fit, 
                         new_data = new_points,
                        interval = 'confidence')
head(confint2)
tail(confint2)
```
# Model 3. Fitting a logistic model to Nausea using RunnyNose the main predictor of interest
```{r}
# Using the tidymodels package, I am going to set up the logistic model
# Set the engine
lm_mod1 <-  logistic_reg() %>% 
  set_engine("glm") 
# Now I am fitting the model
lm_fit3 <- 
  lm_mod1 %>% 
  fit(Nausea ~ RunnyNose, data = processeddta)
tidy(lm_fit3)
```
### Model as box and whiskers plot
```{r}
tidy(lm_fit3) %>% 
dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))

```

# Model 4. Fitting a logistic model to Nausea using all relevant predictors of interest + runnynose
If we remember from our EDA, the interesting predictors were

- Fatigue
- ChillsSweats (y/n)
- SubjectiveFever (y/n)
- Pharyngitis

```{r}
# Using the tidymodels package, I am going to set up the linear model
# Set the engine
lm_mod1 <-  logistic_reg() %>% 
  set_engine("glm") 
# Now I am fitting the model
lm_fit4 <- 
  lm_mod1 %>% 
  fit(Nausea ~ RunnyNose * ChillsSweats * Fatigue * SubjectiveFever* Pharyngitis, data = processeddta)
tidy(lm_fit4)

```
### Model as box and whiskers plot
```{r}
tidy(lm_fit4) %>% 
dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))

```

# Comparing Model 3 and Model 4
To best compare these models and see if they are significant, lets use an ANOVA table
```{r}
tidy(lm_fit3)
tidy(lm_fit4)

ANOVA2<-anova(lm_fit$fit, lm_fit2$fit, test = "Chisq")
ANOVA2

```

P value is significant, on the more complex model. p<0.001











