---
title: "Exploration"
author: "Sophia Drewry"
date: "10/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script loads the processed data does some exploration to prepare for the formal statistical fitting

### Load needed packages
```{r}
library(dplyr) #for data processing
library(here) #to set paths
library(gtsummary)
library(table1)
library(ggplot2)
library(tidyr)
library(survey)
library(ggthemes)
library(RColorBrewer)
library(reshape2)

# path to data
# note the use of the here() package and not absolute paths
dataSPOT <- here::here("data","processed_data","processeddta.rds")

# load data. 
processeddta <-readRDS(dataSPOT)

# take a look at the data
dplyr::glimpse(processeddta)
```

### End goal: 
- Main continuous outcome of interest = Body temperature
- Main categorical outcome = Nausea
- We want to see if the other symptoms are correlated with (predict) those outcomes
- Not as important now since this script is just EDA, but its important to keep in mind

```{r}
## Table one
# Lets get a good understanding of our variables in table form

# I want to produce a quick table to see the most common symptoms
processeddta %>% tbl_summary() %>% 
  modify_caption("**Table 1. Patient Characteristics**")
```

## Ok the ones that look interesting to me are...
- Fatigue (y/n)
- CoughIntensity (none/mild/moderate/severe)
- ChillsSweats (y/n)
- SwollenLymphNodes (y/n)
- Pharyngitis (y/n)
- Insomnia (y/n)
- Headache (y/n)
- SubjectiveFever (y/n)


# Exploring Nausea Variable
```{r}
# Plotting Nausea Variable
processeddta %>% ggplot(aes(x=Nausea, fill=Nausea)) + 
  geom_bar() +
  labs(title= "Nausea Variable Dist.") +
  theme_wsj() +
  scale_fill_brewer(palette = "Dark2")
```


```{r}
# Lets look at the variable distributions compared to Nausea
processeddta %>% 
  select("Nausea", "Fatigue", "CoughIntensity", "ChillsSweats", "SwollenLymphNodes","Pharyngitis", "Insomnia", "Headache", "SubjectiveFever") %>% 
  tbl_summary(by = Nausea) %>% 
  modify_caption("**Table 2. Patient Characteristics vs. Nausea**")
```
Nausea, Fatigue, Chills Sweats, Pharyngitis, and Headache seem to be the most common among those with Nausea. 
For no we will keep all the variables, lets look at them in bar plots

```{r}
# Now lets separate them
nauseads <- processeddta[, c("Nausea", "Fatigue", "ChillsSweats", "SwollenLymphNodes","Pharyngitis", "Insomnia", "Headache", "SubjectiveFever")]

# Changing the data format so we can graph together
uglytry<-melt(nauseads, id.vars = c("Nausea"))
# Not including "CoughIntensity" because it makes graphing complicated
# Note, the new names are Nausea, variable, value

# Now to graph the frequency of Nausea Cases subset by Symptoms
uglytry %>% ggplot(aes(x=value, fill=Nausea)) + 
  geom_bar() +
  labs(title= "Number of Nausea Cases subset by Symptoms") +
  facet_wrap(~variable) +
  theme_wsj() 
```
This looks interesting, lets see if we can see it as a proportion
```{r}
uglytry %>% ggplot(aes(x=value, fill=Nausea)) + 
  geom_bar(position="fill") +
  labs(title= "Number of Nausea Cases subset by Symptoms") +
  facet_wrap(~variable) +
  theme_wsj() 
```
Not as exciting. Nausea distribution seems to differ slightly in fatigue, chill sweats and headache
Now lets move on to body temperature

# Exploring Body Temperature Variable 
```{r}
# Distribution of body temperature
processeddta %>% ggplot(aes(BodyTemp)) + 
  geom_histogram() +
  labs(title= "Distribution of Body Temperature") +
  theme_wsj() +
  scale_fill_brewer(palette = "Dark2")
```
It looks like a majority have normal body temperatures, but there are a good bit outside the normal body temp range. I would be curious to look at fever temperatures. See later code to look at predictors on fever temperature.

# Predictor variables and Body Temp
```{r}
# Cough Intensity
processeddta %>% ggplot( mapping = aes(x = CoughIntensity, y = BodyTemp)) +
  geom_boxplot(aes(fill = CoughIntensity))+ 
  labs(title= "Cough Intensity vs. BodyTemp > 100.4") +
  theme_wsj() +
  scale_fill_brewer(palette = "Dark2")
```
Nothing too interesting here, they all appear to be similar
```{r}
  # Fatigue
processeddta %>% ggplot( mapping = aes(x = Fatigue, y = BodyTemp)) +
  geom_boxplot(aes(fill = Fatigue))+ 
  labs(title= "Fatigue vs. BodyTemp") +
  theme_wsj() +
  scale_fill_brewer(palette = "Dark2")


  # ChillSweats
processeddta %>% ggplot( mapping = aes(x = ChillsSweats, y = BodyTemp)) +
  geom_boxplot(aes(fill = ChillsSweats))+ 
  labs(title= "ChillsSweats vs. BodyTemp") +
  theme_wsj() +
  scale_fill_brewer(palette = "Dark2")

```
Both Fatigue and Chill Sweats seem to follow a similar pattern. May be worth putting in a model.
```{r}
  # SubjectiveFever
processeddta %>% ggplot( mapping = aes(x = SubjectiveFever, y = BodyTemp)) +
  geom_boxplot(aes(fill = SubjectiveFever))+ 
  labs(title= "SubjectiveFever vs. BodyTemp") +
  theme_wsj() +
  scale_fill_brewer(palette = "Dark2")
```
Subjective fever appears to be an interesting figure as well.
```{r}
  #Pharyngitis
processeddta %>% ggplot( mapping = aes(x = Pharyngitis, y = BodyTemp)) +
  geom_boxplot(aes(fill = Pharyngitis))+ 
  labs(title= "Pharyngitis vs. BodyTemp") +
  theme_wsj() +
  scale_fill_brewer(palette = "Dark2")
  
```

Pharyngitis appears to not have much of a difference. It does however have a lot of outlier temperatures. I would be interested to see how it compares with those who have a fever body temp (>= 100.4)

# Ok lets look temperatures >= 100.4
I will be sticking with same variables used earlier. 
```{r}

# Ok lets look at fever only
feverds <- processeddta[processeddta$BodyTemp >= 100.4, c("Fatigue", "ChillsSweats", "SwollenLymphNodes", "Pharyngitis", "Insomnia", "Headache", "SubjectiveFever", "CoughIntensity", "BodyTemp", "RunnyNose")]

# Now lets look at some variables versus fever body temperature
feverds %>% tbl_summary() %>% 
  modify_caption("**Table 3. Patient Characteristics vs. BodyTemp > 100.4**")
```
Fatigue, Chillsweats, Pharyngitis, Headache and SubjectiveFever all seem interesting.
Ok, lets see those in a boxplot
```{r}
# Cough Intensity
feverds %>% ggplot( mapping = aes(x = CoughIntensity, y = BodyTemp)) +
  geom_boxplot(aes(fill = CoughIntensity))+ 
  labs(title= "Cough Intensity vs. BodyTemp > 100.4") +
  theme_wsj() +
  scale_fill_brewer(palette = "Dark2")
```
It seems those who reported lack of cough have a higher temperature out of those who have a fever
```{r}
  # Fatigue
feverds %>% ggplot( mapping = aes(x = Fatigue, y = BodyTemp)) +
  geom_boxplot(aes(fill = Fatigue))+ 
  labs(title= "Fatigue vs. BodyTemp > 100.4") +
  theme_wsj() +
  scale_fill_brewer(palette = "Dark2")


  # ChillSweats
feverds %>% ggplot( mapping = aes(x = ChillsSweats, y = BodyTemp)) +
  geom_boxplot(aes(fill = ChillsSweats))+ 
  labs(title= "ChillsSweats vs. BodyTemp > 100.4") +
  theme_wsj() +
  scale_fill_brewer(palette = "Dark2")

```
Both Fatigue and Chill Sweats seem to follow a similar pattern. 
```{r}
  # SubjectiveFever
feverds %>% ggplot( mapping = aes(x = SubjectiveFever, y = BodyTemp)) +
  geom_boxplot(aes(fill = SubjectiveFever))+ 
  labs(title= "SubjectiveFever vs. BodyTemp > 100.4") +
  theme_wsj() +
  scale_fill_brewer(palette = "Dark2")


  #Pharyngitis
feverds %>% ggplot( mapping = aes(x = Pharyngitis, y = BodyTemp)) +
  geom_boxplot(aes(fill = Pharyngitis))+ 
  labs(title= "Pharyngitis vs. BodyTemp > 100.4") +
  theme_wsj() +
  scale_fill_brewer(palette = "Dark2")
```


These two seem to be less likely to have higher temperatures. 
I would have though that those who reported a fever would be more likely to actually have a higher body temperature

For body temperature and specifically fever, I think Fatigue and Chill Sweats show the strongest association with higher temperature in those who have a temperature > 100.4

# New Data
After looking at all the choosed variables, I will put them into one final dataset before model fitting.
As a recap:
- Main continuous outcome of interest = Body temperature
- Main categorical outcome = Nausea

Other predictor variables
- Fatigue (y/n)
- CoughIntensity (none/mild/moderate/severe)
- ChillsSweats (y/n)
- SwollenLymphNodes (y/n)
- Pharyngitis (y/n)
- Insomnia (y/n)
- Headache (y/n)
- SubjectiveFever (y/n)
- Runny nose

```{r}

smalldta <- processeddta[, c("Fatigue", "ChillsSweats", "SwollenLymphNodes", "Pharyngitis", "Insomnia", "Headache", "SubjectiveFever", "CoughIntensity", "BodyTemp", "Nausea", "RunnyNose")]
save_data_location <- here::here("data","processed_data","smalldta.rds")
saveRDS(smalldta, file = save_data_location)

# For fun, I am saving fever DS
save_data_location2 <- here::here("data","processed_data","feverdta.rds")
saveRDS(feverds, file = save_data_location2)
```
